---
title: "Measles Surveillance Analysis: Hamburg, Germany"
author: "Surveillance Data Analysis Team"
date: "`r Sys.Date()`"
format:
  html:
    theme: cosmo
    toc: true
    code-fold: true
---

::: {.callout-note}
This report was generated using artificial intelligence (Claude from Anthropic) under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.

<!--
To indicate human review: Delete the line above about contents not being reviewed, and replace this comment with:
The contents have been reviewed and validated by [Your Name], [Your Role] on [Date].
-->
:::

## Introduction

This report analyzes measles surveillance data for Hamburg, Germany, using data from the `surveillance` package in R. The analysis focuses on temporal patterns and outbreak characteristics to inform public health monitoring and response strategies.

```{r setup}
#| message: false
#| warning: false

library(surveillance)
library(tidyverse)

# Load the data
data("measlesDE")

# Create Hamburg-specific dataset
hamburg_measles <- data.frame(
  time = 1:nrow(measlesDE@observed),
  cases = measlesDE@observed[, "Hamburg"]
)
```

## Overall Surveillance Summary

The dataset covers `r nrow(measlesDE@observed)` weeks of surveillance in Hamburg. Here are the key metrics:

```{r summary-stats}
summary_stats <- data.frame(
  Total_Cases = sum(hamburg_measles$cases),
  Mean_Cases_Per_Week = mean(hamburg_measles$cases),
  Max_Cases = max(hamburg_measles$cases),
  Weeks_With_Cases = sum(hamburg_measles$cases > 0)
)

knitr::kable(summary_stats, 
             caption = "Summary Statistics for Hamburg Measles Cases",
             digits = 2)
```

## Temporal Analysis

### Weekly Pattern
The following visualization shows the weekly pattern of measles cases in Hamburg:

```{r weekly-pattern}
#| fig.cap: "Weekly Measles Cases in Hamburg"
#| fig.height: 6

ggplot(hamburg_measles, aes(x = time, y = cases)) +
  geom_line(color = "darkred") +
  geom_point(color = "darkred", alpha = 0.5) +
  theme_minimal() +
  labs(x = "Time (weeks from start)",
       y = "Number of Cases") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Monthly Aggregation
To better understand the clustering of cases, we aggregate the data monthly:

```{r monthly-pattern}
#| fig.cap: "Monthly Aggregation of Measles Cases in Hamburg"
#| fig.height: 6

# Create monthly aggregation
hamburg_measles$month <- cut(hamburg_measles$time, 
                           breaks = seq(0, 156, by = 4), 
                           labels = 1:39)

monthly_cases <- hamburg_measles %>%
  group_by(month) %>%
  summarize(total_cases = sum(cases))

ggplot(monthly_cases, aes(x = as.numeric(month), y = total_cases)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(x = "Month (from start)",
       y = "Total Cases per Month") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Key Findings

1. **Occurrence Pattern**:
   - Cases were reported in `r summary_stats$Weeks_With_Cases` out of `r nrow(hamburg_measles)` weeks
   - This represents `r round(summary_stats$Weeks_With_Cases/nrow(hamburg_measles) * 100, 1)`% of weeks having at least one case
   - During weeks with cases, the average number was `r round(summary_stats$Total_Cases/summary_stats$Weeks_With_Cases, 2)` cases

2. **Case Distribution**:
   - Total of `r summary_stats$Total_Cases` cases over the surveillance period
   - Maximum of `r summary_stats$Max_Cases` cases in a single week
   - Overall weekly average of `r round(summary_stats$Mean_Cases_Per_Week, 3)` cases

3. **Temporal Characteristics**:
   - Cases appear sporadically rather than in consistent seasonal patterns
   - Several small clusters of cases are visible
   - Long periods of zero cases between occurrences

## Public Health Implications

The data suggests several important implications for public health monitoring and response:

1. **Sporadic Nature**: The sporadic pattern of cases suggests that Hamburg has generally good measles control, with no sustained transmission.

2. **Cluster Response**: When cases do occur, they tend to appear in small clusters, indicating potential for targeted intervention strategies.

3. **Surveillance Effectiveness**: The ability to detect single cases suggests a sensitive surveillance system.

## Limitations

1. This analysis is based on reported cases only; actual infection rates may differ.
2. The temporal resolution is weekly, which may mask some patterns in transmission.
3. Spatial distribution within Hamburg is not available in this dataset.

## Recommendations

1. Maintain current surveillance sensitivity
2. Investigate clusters promptly when they occur
3. Consider analyzing contact patterns during weeks with multiple cases
4. Compare patterns with neighboring regions to identify potential transmission links

## Technical Notes

This analysis was performed using R version `r getRversion()` with the `surveillance` package. The data represents weekly measles case counts for Hamburg, Germany, as provided in the `measlesDE` dataset.
